---
title: "About kwb.resilience"
author: "Andreas Matzinger"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
## Background

kwb.resilience allows calculating different resilience indices for time series.
Equations are described and discussed in the freely available conference paper by [Matzinger et al. (2018)](https://www.researchgate.net/publication/326040304_Quantitative_Beschreibung_der_Resilienz_urbaner_Wassersysteme) (unfortunately in German only). The sample data for the first example in the paper is also contained in the package as a data.frame "oxygen".

The package should make a starting point for a broader discussion of approaches to tackle resilience of urban water infrastructure for applied cases. Accordingly, adaptations and suggestions are encouraged!


## Acknowledgements

The first version of kwb.resilience was developped at [Kompetenzzentrum Wasser Berlin](http://www.kompetenz-wasser.de) in the framework of the project [netWORKS 4](http://www.kompetenz-wasser.de/en/project/networks4/), financed by the German Ministry of Education and Research in the program [FONA](http://www.fona.de). 



## How to use kwb.resilience

Install the package using devtools as follows. For more information check documentation of single functions and the [tutorial](articles/tutorial.html). 

```{r eval = FALSE}
# install.packages("devtools")

devtools::install_github("kwb-r/kwb.resilience", dependencies = TRUE, build_vignettes = TRUE)
```

note: this will also install two handy KWB packages kwb.event and kwb.plot and their CRAN dependencies.



**Reference:**
Matzinger, A., M. Zamzow, M. Riechel, E. Pawlowsky-Reusing, P. Rouault (2018) _Quantitative Beschreibung der Resilienz urbaner Wassersysteme. In:_ Schriftenreihe wasser infrastruktur ressourcen. Vol. 1. 119-127. TU Kaiserslautern. 

